---
title: "R Notebook"
output: html_notebook
---

Import the data and split in train-set and test-set

```{r import_data}
library(haven)
data <- read_dta("data/zambia_height92.dta")
head(data)


#split data in train and test set
set.seed(123)
train <- sample(1:nrow(data), 0.85*nrow(data))

test <- setdiff(1:nrow(data), train)

train_data <- data[train,]
test_data <- data[test,]
```

Check for missing values (SPOILER: NO MISSING VALUES)

```{r missing_values}
sum(is.na(train_data))

```

Cast of variables' type

```{r cast}
#check if all zscore values are integers
sum(train_data$zscore - as.integer(train_data$zscore))

#cast of zscore to integer
train_data$zscore <- as.integer(train_data$zscore)

#factorize gender
train_data$gender <- as.factor(train_data$gender)

#cast c_breastf as integer
train_data$c_breastf <- as.integer(train_data$c_breastf)

#cast c_age as integer
train_data$c_age <- as.integer(train_data$c_age)

#NOTE: to have the month of birth in m_agebirth calculate: decimal_part * 12

#factorize m_education
train_data$m_education <- as.factor(train_data$m_education)

#factorize m_work
train_data$m_work <- as.factor(train_data$m_work)

#factorize region
train_data$region <- as.factor(train_data$region)

#factorize district
train_data$district <- as.factor(train_data$district)

```

Now let's dive into each variable.

First we plot the distribution of the zscore variable.

```{r zscore}
hist(train_data$zscore, breaks = 20, col = "lightblue", main = "Histogram of zscore", xlab = "zscore", freq = FALSE)
#add the normal curve with mean and sd of zscore
curve(dnorm(x, mean = mean(train_data$zscore), sd = sd(train_data$zscore)), col = "red", lwd = 2, add = TRUE)
#add the vertical dotted line at the mean
abline(v = mean(train_data$zscore), col = "blue", lwd = 2, lty = 2)
# Write the median and mean values to check if they're similar 
text(350, 0.003, paste("Median:", round(median(train_data$zscore), 2), "\nMean:", round(mean(train_data$zscore), 2)), cex = 1.5)

```

Notice how this variable seems to follow a normal distribution. Also the mean (and then the median) is significally lower than 0: this is a measure of how the american population is healthier.

Now we study the gender variable
```{r gender}
#plot the barplot of c_gender

```

